Enforcing Boundaries: China's Overseas NGO Law and Operational Constraints for Global Civil Society

### 2023-12-07 README.qmd
Add abstract

China’s 2017 Overseas NGO (ONGO) Law is part of a larger global trend of legal restrictions on international NGOs (INGOs). However, the effects of this global crackdown on INGOs have been difficult to measure since there is often divergence between the formal de jure regulations and the de facto implementation of these laws. The enforcement of these laws also depends on the services provided by these organizations. In particular, authoritarian regimes display conflicting attitudes towards claim-making (e.g. advocacy) INGOs versus service-providing (e.g. humanitarian assistance) INGOs. In this paper, we measure the effect of China’s ONGO law on INGO operations since 2017. We test how INGO operational flexibility under the ONGO law is influenced by organizations’ issue areas. We argue that service provision INGOs have greater flexibility and are allowed to work in a larger number of provinces than their claim-making counterparts. China provides an excellent setting for testing our claim—the 2017 ONGO law regulates both claim-making and service-providing NGOs, and ambiguity in the law gives ample room for arbitrary discretion by authorities. We test our argument using multilevel Bayesian models with administrative data from 635 registered representative offices of INGOs registered in China between 2017–2021. Our results speak to the broader literature on closing civic space and provide an empirical illustration of the practical effect of NGO restrictions on global civil society.

>>>>

China's 2016 Overseas NGO (ONGO) Law is part of a larger global trend of increased legal restrictions on international nongovernmental organizations (INGOs). A growing body of research analyzes the broad effects of this crackdown on INGOs, finding a divergence in formal *de jure* laws and the *de facto* implementation of those laws. The causes and mechanisms of this divergence remains less explored. Why do authoritarian governments allow—and often collaborate—with some INGOs while harshly regulating or expelling others? What determines the openness of the practical legal operating environment for INGOs? In this paper, we use the case of China to explore how political demands to both restrict and embrace INGOs have shaped the international nonprofit sector in the five years since the ONGO Law came into effect. We argue that in an effort to bolster regime stability, governments use civil society laws as policy tools to influence INGO behavior. We find that INGO issue areas, missions, and pre-existing relationships with local government officials influence the degree of operating space available for INGOs. We test this argument with a mixed methods research design, combining Bayesian analysis of administrative data from all formally registered INGOs with a comparative case study of two environmental INGOs. Our findings offer insights into the practical effects of INGO restrictions and the dynamics of closing civic space worldwide.

### 2023-12-07 index.qmd
Add abstract

Abstract goes here

>>>>

China's 2016 Overseas NGO (ONGO) Law is part of a larger global trend of increased legal restrictions on international nongovernmental organizations (INGOs). A growing body of research analyzes the broad effects of this crackdown on INGOs, finding a divergence in formal *de jure* laws and the *de facto* implementation of those laws. The causes and mechanisms of this divergence remains less explored. Why do authoritarian governments allow—and often collaborate—with some INGOs while harshly regulating or expelling others? What determines the openness of the practical legal operating environment for INGOs? In this paper, we use the case of China to explore how political demands to both restrict and embrace INGOs have shaped the international nonprofit sector in the five years since the ONGO Law came into effect. We argue that in an effort to bolster regime stability, governments use civil society laws as policy tools to influence INGO behavior. We find that INGO issue areas, missions, and pre-existing relationships with local government officials influence the degree of operating space available for INGOs. We test this argument with a mixed methods research design, combining Bayesian analysis of administrative data from all formally registered INGOs with a comparative case study of two environmental INGOs. Our findings offer insights into the practical effects of INGO restrictions and the dynamics of closing civic space worldwide.

### 2023-12-07 manuscript/manuscript.qmd
Add abstract

---
title: "Enforcing Boundaries: China's Overseas NGO Law and Operational Constraints for Global Civil Society"
short-title: "Enforcing Boundaries"
date: "2023-12-07"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  TODO: **Update this based on the finished version of the paper.** China’s 2016 Overseas NGO (ONGO) Law is part of a larger global trend of legal restrictions on international nongovernmental organizations (INGOs). However, the effects of this global crackdown on INGOs have been difficult to measure since there is often divergence between the formal de jure regulations and the de facto implementation of these laws. Additionally, the severity and consistency of the enforcement of these laws often depends on the type of services provided by these organizations. In this paper, we explore the effect of China’s ONGO law on INGO operations in the five years since the passage of the law. We test whether and how INGO operational space under the ONGO law is influenced by organizations' issue areas. We argue that the less contentious the issue areas INGOs work in the more likely they are granted greater operational space, measured by the number of provinces they are legally authorized to work in. China provides an excellent setting for examining the phenomenon of closing space civic space, since the 2016 ONGO law hypothetically regulates both claim-making and service-providing NGOs, while the level of ambiguity of the law gives ample room for arbitrary discretion and conflicting attitudes by authorities. We test our argument by analyzing administrative data from 633 registered representative offices of INGOs registered in China between 2017–2021 and test our hypothesis with a multilevel Bayesian model. Our results speak to the broader literature on closing civic space and provide an empirical illustration of the practical effect of NGO restrictions on global civil society.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
word-count: 7984
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
toc-depth: 4
---

>>>>

---
title: "Enforcing Boundaries: China's Overseas NGO Law and Operational Constraints for Global Civil Society"
short-title: "Enforcing Boundaries"
date: "2023-12-07"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  China's 2016 Overseas NGO (ONGO) Law is part of a larger global trend of increased legal restrictions on international nongovernmental organizations (INGOs). A growing body of research analyzes the broad effects of this crackdown on INGOs, finding a divergence in formal *de jure* laws and the *de facto* implementation of those laws. The causes and mechanisms of this divergence remains less explored. Why do authoritarian governments allow—and often collaborate—with some INGOs while harshly regulating or expelling others? What determines the openness of the practical legal operating environment for INGOs? In this paper, we use the case of China to explore how political demands to both restrict and embrace INGOs have shaped the international nonprofit sector in the five years since the ONGO Law came into effect. We argue that in an effort to bolster regime stability, governments use civil society laws as policy tools to influence INGO behavior. We find that INGO issue areas, missions, and pre-existing relationships with local government officials influence the degree of operating space available for INGOs. We test this argument with a mixed methods research design, combining Bayesian analysis of administrative data from all formally registered INGOs with a comparative case study of two environmental INGOs. Our findings offer insights into the practical effects of INGO restrictions and the dynamics of closing civic space worldwide.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
word-count: 7984
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
toc-depth: 4
---
